\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-29 paper/paper.tex
Clean up semantics in the paper.

A key value store typically serves multiple client requests concurrently and is
linearizable~\cite{linearizability}. This means any history of requests and
responses can be rearranged such that responses immediately follow corresponding
requests and if a response precedes a request in the original history it also
does so in the rearranged history. Key value stores such as RocksDB also
support atomic writes ({\tt PUT} operations) to multiple keys, called batched
writes, and the ability to take snapshots.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs in sync
across the replicas. On failure of the leader replica, a new leader is
automatically elected -- RAFT ensures that the new leader has the most up to
date log. The Cyclone client library automatically locates the new leader on a
failover, resulting in only a brief interruption in service for
clients. Although read requests (GET operations) are not logged, they are
serviced in FIFO order with the write requests (PUT/DELETE operations) in
Cyclone's log at the moment the read request is received. This ensures
linearizability in the face of failovers.

@@

The first question we deal with is -- how do Cyclone clients decide which
physical log to send a request to? A key constraint here is that all replicas
should agree on the sequence of updates to a key, as otherwise they would
diverge in terms of state. Allowing updates to the same key to go to different
physical logs would lead to a divergence as there is no guarantee about the
order in which the independent RAFT instances would apply the update. Therefore
all updates to a key must go to the same physical log. A similar constraint
applies to reads in order to ensure that all committed writes are seen by
readers. Allowing reads to a key to go to a different RAFT instance than the
write means that it could be served on a machine where the write is yet to be
applied. Therefore, all reads and writes to a key must go to the same physical
log. We achieve this by hashing the key to select the physical log.

It turns out that this simple \emph{necessary} scheme is also \emph{sufficient}
to ensure linearizability in the face of failure. Clearly the system is
linearizable for a single key since it involves only a single physical log an d
we always service existing requests in the physical log before responding to
read requests. In addition, one can switch operations to different keys in a
history without affecting its correctness because operations to different keys
are commutative. This allows us to arrive at a history that explains the key
value store responses, is correct and makes sure that a request responded to
before subsequent requests in the original history follows that same order in
the linearized history. We discuss how to handle requests that atomically
manipulate multiple keys in the next section.

We note that spreading keys across the physical logs \emph{does not} break
referential integrity. If key1 is updated followed by key2 and key2 refers to
key1, even in the face of failure linearizability in Cyclone ensures that key1
is available once key2 is seen by a client. This is true even if key1 is not
committed yet at the leader replica, since we execute reads in FIFO order with
pending entries in the log (Section~\ref{sec:sysarch}).

@@

In our discussion thus far, we have focused on operations to a single
key. Supporting atomic writes or reads to multiple keys while maintaining
linearizability requires special handling with multiple physical logs so as not
to break linearizability - that we discuss next.

\subsection{Ganged Operations}
Supporting reads or writes to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together even in the face of failures. One option is to have an
external co-ordinator run a distributed transaction using two phase commit. In
addition to requiring extra external resources to host the co-ordinator one must
also worry about failure recovery. This also requires two round trips to the
leader to complete what should have been a single batched write request from a
client that is stateless and allowed to fail - such as with key value stores
like Rocksdb~\cite{rocksdb}. We therefore describe a simpler and more performant
solution called ``ganged operations" in Cyclone. Ganged operations allow a
stateless client to issue a batch of update operations as a ``gang",
simultaneously and atomically updating a batch of keys in a single call.

>>>>

Key value stores such as RocksDB also support atomic writes ({\tt PUT}
operations) to multiple keys, called batched writes, and the ability to take
snapshots.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs in sync
across the replicas. On failure of the leader replica, a new leader is
automatically elected -- RAFT ensures that the new leader has the most up to
date log. The Cyclone client library automatically locates the new leader on a
failover, resulting in only a brief interruption in service for
clients. Although read requests (GET operations) are not logged, they are
serviced in FIFO order with the write requests (PUT/DELETE operations) in
Cyclone's log at the moment the read request is received.

Cyclone provides the following guarantees to clients:

\begin{itemize}
\item Every acknowledged update is applied
\item Every key sees the same sequence of updates across all replicas
\item A client sees a monotonically extending prefix of the sequence of updates
  to any key at that always includes its own updates
\end{itemize}

RAFT ensures that the set of committed updates to any key is in the same order
as in the log, satisfying the first two conditions. For the last condition,
clients maintain the term of the RAFT leader locally and ensure they always talk
to a (possibley different) RAFT leader that has at least the same term. This
helps us avoid split brain problems where a network partitions allows clients to
talk to old RAFT leaders after having committed changes to a new RAFT leader.

We do not provide linearizable reads, although that can be easily supported by
logging the read and running conensus on it like a regular mutation operation.
Finally, Cyclone allows operations to multiple keys in an atomic batch, where
either all updates in the batch are applied or no operation in the batch is
applied, regardless of failures. We also use a lightweight batched operation to
support snapshots.

@@

The first question we deal with is -- how do Cyclone clients decide which
physical log to send a request to? The guarantees from Cyclone cover ordering of
updates to a single key (Section~\ref{sec:sysarch}). They can be satisfied by
ensuring that all reads and writes to a key go to the same physical log
(i.e. RAFT instance). We achieve this by hashing the key to select the physical
log.

@@

\subsection{Ganged Operations}
Supporting batched updates to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together even in the face of failures. One option is to have an
external co-ordinator run a distributed transaction using two phase commit. In
addition to requiring extra external resources to host the co-ordinator one must
also worry about failure recovery. This also requires two round trips to the
leader to complete what should have been a single batched write request from a
client that is stateless and allowed to fail - such as with key value stores
like Rocksdb~\cite{rocksdb}. We therefore describe a simpler and more performant
solution called ``ganged operations" in Cyclone. Ganged operations allow a
stateless client to issue a batch of update operations as a ``gang",
simultaneously and atomically updating a batch of keys in a single call.
